Parallel Hierarchical Clustering using \\ Rank-Two Nonnegative Matrix Factorization \thanks{Identify applicable funding agency here. If none, delete this.}

### 2020-07-16 latex/tex/experiments.tex
edited hier perf section

Figure \ref{fig:synhierstrongscaling} shows the limits of scaling as the hierarchical cost is becoming saturated with communication.
The AllReduce segment would consistently grow if the problem was becoming latency bound, which is not the case.
So, the problem is becoming bandwidth bound as seen through the ReduceScatter and AllGather segments. 
However, computation still makes up more than 60\% of the total cost.

With the much larger \image{} dataset, it's possible to scale much further as seen in Figure \ref{fig:rwhierspeedup} without being overwhelmed by the communication cost.
From Figure \ref{fig:fig:rwhierstrongscaling}, the communication cost doesn't make up more than 20\% of the total cost even scaling to 80 compute nodes.

Note that in order to compare hierarchical clusters between differing compute nodes, we limited the output to include only complete levels in the tree.

\begin{itemize}
	\item theory (\Cref{eq:treecost}) says...
	\item Fig 7 shows limit of strong scaling for synthetic data (time increases at 50 nodes compared to 40), Fig 8 gives breakdown and explains
	\begin{itemize}
		\item drastic performance dropoff may be artifact of all-gather, which should take about the same time as reduce-scatter, but speedup from 30 to 40 nodes is also slight
		\item by 50 nodes, about 50\% of time in communication; still mostly dominated by bandwidth cost because ag and rs dominate ar
		\item note the difference between the scaling of single rank-2 and entire tree: poor scaling must be due to lower levels, which we explore in the next section
	\end{itemize}
	\item Fig 9 shows reasonable scaling for entire tree -- speedup drops from 7x to 6x over 80 nodes; Fig 10 shows that MM is still 70\% of run time on 80 nodes, which allows for near-linear scaling
\end{itemize}

>>>>

\Cref{fig:synhierspeedup} demonstrates the scalability of the synthetic data set on up to 40 nodes, and we observe a 15$\times$ speedup compared to 1 node.
\Cref{fig:synhierstrongscaling} shows the relative time breakdown and explains the limitation on scaling.
On 40 nodes, computation still takes 60\% of the total time, but the all-gather and reduce-scatter costs have grown in relative time because they do not scale with $p$.
Because all-reduce involves only a constant amount of data and its time remains relatively small, we conclude the communication is bandwidth bound at this scale.

With the larger \image{} dataset, it's possible to scale much further as seen in \Cref{fig:rwhierspeedup}, where we observe a 5.9$\times$ speedup of 80 compute nodes compared to 10.
From \Cref{fig:fig:rwhierstrongscaling}, we see that the communication cost constitutes less than 20\% of the total time even at 80 compute nodes.

We note that the speedup of the overall hierarchical clustering algorithm is not as high as for a single Rank-2 NMF (measured at the root node).
This is due to inefficiencies in the lower levels of the tree, as we explore in the next section.

%Note that in order to compare hierarchical clusters between differing compute nodes, we limited the output to include only complete levels in the tree.
